---
title: "Visualization and EDA"
author: "Kamiah Brown"
date: "2024-09-26"
output: github_document
---
```{r}
library(tidyverse)
library(ggridges)
library(haven)
```
 

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Making our first plot 
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point( alpha = .3, size = .8)
geom_smooth(se = FALSE)
```

Where you define aesthetics can matter
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(aes(color = name), alpha = .3, size = .8)
geom_smooth(se = FALSE)
```
Use faceting; three different panels by each name
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

interesting scatterplot 
```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name, size = prcp)) +
  geom_point( alpha = .3) + 
  geom_smooth(se = FALSE)
  facet_grid(. ~ name)
```

Learning assessment 
```{r}
weather_df |>
  filter(name == "CentralPark_NY") |>
  mutate(
    tmax_fahr = tmax * (9 / 5) + 32,
    tmin_fahr = tmin * (9 / 5) + 32
  ) |>
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point()
geom_smooth(method = "lm", se = FALSE)
```

Small things 
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .3, size =.8) +
  geom_smooth(se = FALSE) 
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point( color = "blue")
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = "blue")) +
  geom_point()
```

If something is a variable name, it doesn't need to be in quotes. If it is not a variable name then it is not a variable name. 

##Univeraite plots 
```{r}
weather_df |>
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

Use fill to fill in the histogram 

```{r}
weather_df |>
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram()
facet_grid(. ~ name)
```

density plot 
```{r}
weather_df |>
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .3)
```
Box plots
```{r}
weather_df |>
  ggplot(aes(x = name, y = tmin, fill = name)) +
geom_boxplot()
```
Violin plots
```{r}
weather_df |>
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin()
```
Ridge plots 
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```

Learning assessment comparing prepipation across locations
```{r}
weather_df |>
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 3)

weather_df |>
  ggplot(aes(x = name, y = prcp)) +
  geom_boxplot()

weather_df |>
  filter(prcp > 10, prcp < 1000) |>
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .3)
```

## Saving and embedding plots 
```{r}
ggp_weather =
  weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()

ggsave("ggp_weather.pdf", ggp_weather, width = 8, height = 6)
```

Embedding plots 
```{r, fig.width = 6, fig,asp = .6}
  weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()
```


Make a scatterplot
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3)
labs(
  title = "Temperature scatterplot",
  x = "Minimum Temp (C)",
  y = "Maximum Temp (C)",
  color = "Location",
  caption = "Weather data taken fron rnoaa package for three stations."
)
```

Scales --
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3)
labs(
  title = "Temperature scatterplot",
  x = "Minimum Temp (C)",
  y = "Maximum Temp (C)",
  color = "Location",
  caption = "Weather data taken fron rnoaa package for three stations."
  ) + 
  scale_x_continuous(
    breaks = c(-15, 0, 20),
    labels = c("-15C", "0", "20")
  ) + 
  scale_y_continuous(
    limits = c(0, 30), 
    transform = "sqrt"
  )
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3)
labs(
  title = "Temperature scatterplot",
  x = "Minimum Temp (C)",
  y = "Maximum Temp (C)",
  color = "Location",
  caption = "Weather data taken fron rnoaa package for three stations."
) 
viridis::scale_color_viridis(discrete = TRUE) + 
  theme(legend.position = "bottom") +
  theme_bw()
```


#Themes
```{r}
gg_scatterplot +
  theme(legend.position = "bottom")
```

```{r}
ggp_scatterplot + 
  theme_bw() +
  theme(legend.position = "bottom")
```

```{r}
ggp_scatterplot + 
  theme(legend,postion = "bottom")
  theme_minimal()
```

```{r}
gg_scatterplot +
  theme(legend.positon)
```

```{r}

```


Learning Assessment 
```{r}
weather_df  |>
  ggplot(aes( x = date, y = tmax, color = name, size = prcp)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) + 
       viridis::scale_color_viridis(discrete = TRUE) +
       labs( 
         x = "Date",
         y = "Maximum Temp (C)",
         title = "Seasonal variaiton in Max Temp"
         ) + 
         theme_minimal() + 
         theme(legend.position = "bottom")
```



Extra bonus stuff in 'ggplot' 
Use different datasets in different 'geom's
```{r}
central_park_df = 
  weather_df |> 
  filter(name == "CentralPark_NY")

molokai_df = 
weather_df |>
  filter(name == "Molokai_HI")

molokai_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park_df)
```

#Multiple panels
```{r}
weather_df |>
  ggplot(aes(x = tmax, fill = name)) +
  geom_density() +
  facet_grid(. ~ name)
```

```{r}
ggp_tmax_tmin = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3)

ggp_tmax_density = 
  weather_df |>
  ggplot(aes(x = timax, fill = name)) +
  geom_density(alpha = .3)

ggp_tmax_date =
  weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()
geom_smooth(se = FALSE)
```

#Data maniupulation 
```{r}
weather_df |>
  mutate(name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", 
                                    "Waterhole_WA"))) |>
  ggplot(aes(x = name, y = tmax, fill = name)) + 
  geom_violin(alpha = .5)
```
Pulse data
```{r}
pulse_data = 
 read_sas("/data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer(
    cols = bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    values_to = "bdi_score", 
    names_prefix_ = "bdi_score_"
  )|>
  
  ggplot(pulse_data, aes(x = visit, y = bdi)) + 
  geom_boxplot()
```


Make an FAS plot 
```{r}
litters_df = 
  read_csv("FAS_litters.csv", na = c("NA", "." "")) |>
  janitor::clean_names() |>
  separate(group, into = c("dose", "tx_day"), 3)

pups_df = 
  read_csv("FAS_pups.csv", na =c("NA", "." "")) |>
  janitor::clean_names() |>
  
  
```

```{r}
weather_df |>
  filter(name == "CentralPark_NY") |>
  mutate(
    tmax_fahr = tmax * (9 / 5) + 32,
    tmin_fahr = tmin * (9 / 5) + 32
  ) |>
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point()
geom_smooth(method = "lm", se = FALSE)
```


```{r}
weather_df |>
  ggplot(aes(x = prcp))
geom_histogram()
```

```{r}
weather_df |>
  filter(prcp > 1000)
```

```{r}
weather_df |>
  filter()
ggplot(aes(x = tmin, y = tmax, color = name, sjape = name)) +
  geom_point()
```

### Group_by
missing month data. 
```{r}
weather_df |>
  group_by (name)
```


#counting stuff 
```{r}
weather_df |> 
  group_by (name) |>
  summarize(n_obs = n(), 
            n_dist = n_distinct(name))
```

#2x2

```{r}
weather_df |>
  drop_na(tmax) |>
  filter( name != "Molokai_HI") |>
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not_cold"
    )
  ) |> 
  group_by(name, cold) |>
  summarize(count = n())
```

```{r}
weather_df |>
  drop_na(tmax) |>
  filter( name != "Molokai_HI") |>
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not_cold"
    )
  ) |> 
    janitor::tabyl(name, cold)
```

## generic numeric summaries
 lets try some other useful summaries 
 
```{r}
weather_df |>
  group_by(name) |>
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE),
    sd_prcp = sd(prcp, na.rm = TRUE)
  )
```

summarize and then plot 
```{r}
weather_df |>
  group_by(name) |>
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE),
    sd_prcp = sd(prcp, na.rm = TRUE)
  )

#Missing month.
ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_line()
```


