---
title: "Visualization and EDA"
author: "Kamiah Brown"
date: "2024-09-26"
output: github_document
---
```{r}
library(tidyverse)
library(ggridges)
```
 

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Making our first plot 
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point( alpha = .3, size = .8)
geom_smooth(se = FALSE)
```

Where you define aesthetics can matter
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(aes(color = name), alpha = .3, size = .8)
geom_smooth(se = FALSE)
```
Use faceting; three different panels by each name
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

interesting scatterplot 
```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name, size = prcp)) +
  geom_point( alpha = .3) + 
  geom_smooth(se = FALSE)
  facet_grid(. ~ name)
```

Learning assessment 
```{r}
weather_df |>
  filter(name == "CentralPark_NY") |>
  mutate(
    tmax_fahr = tmax * (9 / 5) + 32,
    tmin_fahr = tmin * (9 / 5) + 32
  ) |>
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point()
geom_smooth(method = "lm", se = FALSE)
```

Small things 
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .3, size =.8) +
  geom_smooth(se = FALSE) 
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point( color = "blue")
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = "blue")) +
  geom_point()
```

If something is a variable name, it doesn't need to be in quotes. If it is not a variable name then it is not a variable name. 

##Univeraite plots 
```{r}
weather_df |>
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

Use fill to fill in the histogram 

```{r}
weather_df |>
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram()
facet_grid(. ~ name)
```

density plot 
```{r}
weather_df |>
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .3)
```
Box plots
```{r}
weather_df |>
  ggplot(aes(x = name, y = tmin, fill = name)) +
geom_boxplot()
```
Violin plots
```{r}
weather_df |>
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin()
```
Ridge plots 
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```

Learning assessment comparing prepipation across locations
```{r}
weather_df |>
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 3)

weather_df |>
  ggplot(aes(x = name, y = prcp)) +
  geom_boxplot()

weather_df |>
  filter(prcp > 10, prcp < 1000) |>
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .3)
```

## Saving and embedding plots 
```{r}
ggp_weather =
  weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()

ggsave("ggp_weather.pdf", ggp_weather, width = 8, height = 6)
```

Embedding plots 
```{r, fig.width = 6, fig,asp = .6}
  weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()
```

